A boundary-integral framework to simulate \\ viscous erosion of a porous medium

### 2018-04-16 docs/methods/erosion_methods.tex
NM

To numerically compute the completed double-layer potential~\eqref{eqn:completed_BIE}, we discretize each of the $M$ bodies with  $N_\iin$ points, and the outer boundary with $N_\out$ points.  Then, $N = MN_\iin + N_\out$ is the total number of discretization points of $\bd\Omega$.  The double-layer potential in~\eqref{eqn:DLP} is discretized with the trapezoid rule as
\begin{align}
  \DD[\eeta](\xx_i) = \sum_{j=1}^{N} K(\xx_i,\xx_j) \eeta(\xx_j) 
      \Delta s_{j}\, , \quad i=1,\ldots,N \, ,
  \label{eqn:trapDLP}
\end{align}
where $\Delta s_j$ is the arclength term of $\bd\Omega$ at
$\xx_j$, and
\begin{align}
  K(\xx,\yy) = \frac{1}{\pi} \frac{\rr \cdot \nn}{\rho^2} 
      \frac{\rr \otimes \rr}{\rho^2} \, ,
\end{align}
where $\rr = \xx - \yy$ and $\rho = \|\rr\|$, is the kernel of the double-layer potential.  The diagonal term of~\eqref{eqn:trapDLP} is replaced with the limiting value of the double-layer potential
\begin{align}
  D(\xx,\xx) = -\frac{\kappa(\xx)}{2\pi}(\ss(\xx) \otimes \ss(\xx)) \, ,
\end{align}
where $\kappa(\xx)$ is the curvature and $\ss(\xx)$ is the unit tangent vector at $\xx \in \bd\Omega$.  The equations for the Stokeslets~\eqref{eqn:stokeslet} and the rotlets~\eqref{eqn:rotlet} are also discretized with the trapezoid rule.  Applying this discretization strategy to~\eqref{eqn:completed_BIE}, the dense linear system to be solved is
%\begin{equation}
%\todo[inline]{Typo above}
%\begin{equation}
\begin{subequations}
  \label{eqn:completed_BIE_discretized}
  \begin{align}
  \label{eqn:discretized_DLP}
%  \begin{aligned} 
%    \ff(\xx_i) &= -\frac{1}{2}\ssigma(\xx_i) + \sum_{j=1}^{N} 
%      K(\xx_i,\xx_j) \eeta(\xx_j) \Delta s_{j} + 
%      \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx_i) \\
%      &\hspace{53pt}
%      +\sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx_i) 
%      +\sum_{\xx_j \in \Gamma} (\nn(\xx_i) \otimes
%      \nn(\xx_j))\eeta(\xx_j) \, , \\
    \ff(\xx_i) &= -\frac{1}{2}\ssigma(\xx_i) + \sum_{j=1}^{N} 
      K(\xx_i,\xx_j) \eeta(\xx_j) \Delta s_{j} + 
      \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx_i) +
      \sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx_i) \\
    \llambda_\ell &= \sum_{\xx_j \in \gamma_\ell} \eeta(\xx_j) 
      \Delta s_j \, , \\ 
    \xi_\ell &= \sum_{\xx_j \in \gamma_\ell}
      {(\xx_j - \cc_\ell)}^\perp \cdot \eeta(\xx_j) \Delta s_j \, ,
  \end{align}
\end{subequations}
for $i=1,\ldots,N$, $\ell=1,\ldots,M$.  When $\xx_i \in \Gamma$, the term 
\begin{align}
  \sum_{\xx_j \in \Gamma} (\nn(\xx_i) \otimes
      \nn(\xx_j))\eeta(\xx_j) \, ,
\end{align}
is added to~\eqref{eqn:discretized_DLP}.  Since we discetize integrands that are smooth and periodic, the trapezoid rule employed above achieves spectral accuracy~\cite{tre-wei2014}.  
%We have applied the trapezoid rule since all the integrands are smooth and periodic, and this guarantees that the numerical solutions converges to the true solution with spectral accuracy~\cite{tre-wei2014}.
%\todo[inline]{The above sentence strikes me as awkward and verbose. Rephrase? \\
%How about this: Since all integrands are smooth and periodic, the trapezoid rule employed above achieves spectral accuracy.}

@@

\subsection{Convergence test}
As a final validation, we now verify that our method converges with the expected second-order accuracy in time. For this, we run the same single-body simulation shown in Figs.~\ref{01bodseq} and~\ref{shrink_intface} (except with a coarser $N_\iin=256$) to a stopping time of $t_s = 10^{-2}$. This stopping time is 56\% of the vanishing time, and was selected so that the body will have eroded substantially but is not small enough to compromise accuracy (due to the stress diverging).  We assess error by measuring the $L^2$-difference in the body shape obtained from simulations of successive resolution, with $\Delta t$ halved at each stage. In Table~\ref{convtab}, we report these errors along with the order of accuracy, obtained by Richardson-extrapolation-type calculations. The table confirms that our method converges with expected second-order accuracy. The last column of the table shows the CPU time taken for the simulation to run on a single-processor machine, demonstrating a nearly linear increase in computational time. Though the main parts of the algorithm are easily paralyzed, the single-processor times provide a simple assessment of the overall complexity.
\todo[inline]{Aren't the CPU scaling times we report obvious.  If we take a time step size half the size, it will of course take twice as long.  We report these numbers in a couple of tables.  We have not reported CPU time versus the spatial resolution which is where the value of the FMM would be obvious.}

>>>>

To numerically compute the completed double-layer potential~\eqref{eqn:completed_BIE}, we discretize each of the $M$ bodies with  $N_\iin$ points, and the outer boundary with $N_\out$ points.  Then, $N = MN_\iin + N_\out$ is the total number of discretization points of $\bd\Omega$.  The double-layer potential in~\eqref{eqn:DLP} is discretized with the trapezoid rule as
\begin{align}
  \DD[\eeta](\xx_i) = \sum_{j=1}^{N} K(\xx_i,\xx_j) \eeta(\xx_j) 
      \Delta s_{j}\, , \quad i=1,\ldots,N \, ,
  \label{eqn:trapDLP}
\end{align}
where $\Delta s_j$ is the arclength term of $\bd\Omega$ at
$\xx_j$, and
\begin{align}
  K(\xx,\yy) = \frac{1}{\pi} \frac{\rr \cdot \nn}{\rho^2} 
      \frac{\rr \otimes \rr}{\rho^2} \, ,
\end{align}
where $\rr = \xx - \yy$ and $\rho = \|\rr\|$, is the kernel of the double-layer potential.  The diagonal term of~\eqref{eqn:trapDLP} is replaced with the limiting value of the double-layer potential
\begin{align}
  D(\xx,\xx) = -\frac{\kappa(\xx)}{2\pi}(\ss(\xx) \otimes \ss(\xx)) \, ,
\end{align}
where $\kappa(\xx)$ is the curvature and $\ss(\xx)$ is the unit tangent vector at $\xx \in \bd\Omega$.  The equations for the Stokeslets~\eqref{eqn:stokeslet} and the rotlets~\eqref{eqn:rotlet} are also discretized with the trapezoid rule.  Applying this discretization strategy to~\eqref{eqn:completed_BIE}, the dense linear system to be solved is
\begin{subequations}
  \label{eqn:completed_BIE_discretized}
  \begin{align}
  \label{eqn:discretized_DLP}
    \ff(\xx_i) &= -\frac{1}{2}\ssigma(\xx_i) + \sum_{j=1}^{N} 
      K(\xx_i,\xx_j) \eeta(\xx_j) \Delta s_{j} + 
      \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx_i) +
      \sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx_i) \\
    \llambda_\ell &= \sum_{\xx_j \in \gamma_\ell} \eeta(\xx_j) 
      \Delta s_j \, , \\ 
    \xi_\ell &= \sum_{\xx_j \in \gamma_\ell}
      {(\xx_j - \cc_\ell)}^\perp \cdot \eeta(\xx_j) \Delta s_j \, ,
  \end{align}
\end{subequations}
for $i=1,\ldots,N$, $\ell=1,\ldots,M$.  When $\xx_i \in \Gamma$, the term 
\begin{align}
  \sum_{\xx_j \in \Gamma} (\nn(\xx_i) \otimes
      \nn(\xx_j))\eeta(\xx_j) \, ,
\end{align}
is added to~\eqref{eqn:discretized_DLP}.  Since we discetize integrands that are smooth and periodic, the trapezoid rule employed above achieves spectral accuracy~\cite{tre-wei2014}.

@@

\subsection{Convergence test}
As a final validation, we now verify that our method converges with the expected second-order accuracy in time. For this, we run the same single-body simulation shown in Figs.~\ref{01bodseq} and~\ref{shrink_intface} (except with a coarser $N_\iin=256$) to a stopping time of $t_s = 10^{-2}$. This stopping time is 56\% of the vanishing time, and was selected so that the body will have eroded substantially but is not small enough to compromise accuracy (due to the stress diverging).  We assess error by measuring the $L^2$-difference in the body shape obtained from simulations of successive resolution, with $\Delta t$ halved at each stage. In Table~\ref{convtab}, we report these errors along with the order of accuracy, obtained by Richardson-extrapolation-type calculations. The table confirms that our method converges with expected second-order accuracy. The last column of the table shows the CPU time taken for the simulation to run on a single-processor machine.
%, demonstrating a nearly linear increase in computational time with refinement of the time step. 
%Though the main parts of the algorithm are easily paralyzed, the single-processor times provide a simple assessment of the overall complexity.
